---
title: "1. Data Cleaning"
output:
  html_document:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Data Cleaning: dataset 1
```{r}
# Load tidyverse package
library(tidyverse)

# Import our data and save it as datafile
datafile <- read_csv("survey.csv")
```

### head()
```{r}
# View the first five rows with head()
# head(filename, n) shows the first n lines of the data file
head(datafile, 6)
```

### str()
#### the structure of the variables
```{r}
# Check the structure of the variables in the dataset by using str() function
str(datafile)
```

### as.factor()
#### categorical variables
```{r}
# When we check the structure of the data, we see that some of the variables are stored as characters instead of categorical variables. We will set the correct measurement levels for those variables by using as.factor() function.
datafile$Gender <- as.factor(datafile$Gender)
```

### lappy()
```{r}
# In order to change the data types of multiple columns, we will use lapply() function together with as.factor(). lapply() function is very useful for performing operations on list objects such as columns of a dataframe.

# lapply(vector, function)

# Set the correct measurement levels for Gender, Handedness, Exercise and Smoke using lapply() and as.factor() functions

# First generate a vector to keep the column names
columns <- c("Gender", "Handedness", "Exercise", "Smoke")

# Set the correct measurement levels or data types
datafile[columns] <- lapply(datafile[columns], as.factor)
```

### levels()
```{r}
# Check levels of the Gender column
levels(datafile$Gender)
```

### which()
#### find indices
```{r}
# Replace female with Female
# (1) Find the indices of rows with "female" value
( index_female <- which(datafile$Gender == "female") )
```

```{r}
# (2) Replace those entries with "Female"
datafile$Gender[index_female] = "Female"
```

```{r}
# Replace male with Male
# (1) Find the indices of rows with "male" value
( index_male <- which(datafile$Gender == "male") )
```

```{r}
# (2) Replace those entries with "Male"
datafile$Gender[index_male] = "Male"
```

```{r}
# Check levels of the Gender column again
levels(datafile$Gender)
```

```{r}
# Although we corrected the entry errors in the Gender column, the levels did not change. Hence, we should update the factor levels by using factor() function.

# Update the levels
datafile$Gender <- factor(datafile$Gender)
```

### missing value
### summary()
#### statistical information about the data -> check missing values
```{r}
# Check the summary of the data file
summary(datafile)
```

### mean()
```{r}
# Find the average pulse 
mean(datafile$Pulse)
```

### na.omit()
#### remove records with missing values and assign it to `datafile_new`.
```{r}
# Remove records with missing values and assign it to datafile_new
datafile_new <- na.omit(datafile)
```

```{r}
# Calculate the number of records removed from the data
nrow(datafile) - nrow(datafile_new)
```

```{r}
# Find the average of Pulse by excluding the missing values
avg_pulse <- mean(datafile$Pulse, na.rm = T) 
            # na.rm: whether or not to remove NA values from the calculation
print(avg_pulse)
```

```{r}
# Find the average of Height by excluding the missing values
avg_height <- mean(datafile$Height, na.rm = T)

print(avg_height)
```

### replace_na()
```{r}
# Replace records with missing values
datafile_replace <- replace_na(datafile, list(Pulse = avg_pulse, Height = avg_height))

#  Remove records with missing values and assign the redacted dataset to datafile_removed
datafile_removed <- na.omit(datafile_replace)

# Calculate the number of records removed from the data
nrow(datafile) - nrow(datafile_removed)
```

## Data Cleaning: dataset 2
### new dataset
#### `Credit.csv`
```{r}
# Load tidyverse package
library(tidyverse)

# Import the data and save it as crdata
crdata <- read_csv("Credit.csv")
```

```{r}
# View the first four rows with head()
head(crdata, 4)  

# Check the structure of the variables in the dataset by using str() function
str(crdata)

# Check the summary of the dataset
summary(crdata)

# Update the data types if necessary
```

### boxplot()
#### check outliner
```{r}
# outliner in the Age column of crdata
boxplot(crdata$Age)
```

### which()
```{r}
# Find the indices of records with age >= 100 and assign these indices to outliers
outliers <- which(crdata$Age>=100)  

# Print data records with outliers 
print(crdata[outliers,])
```

### remove rows/columns
```{r}
# dataframe[-rowindex,]
```

```{r}
# Remove data records with age >= 100 
crdata <- crdata[-outliers,]

# check the boxplot for Age column again
boxplot(crdata$Age)
```

```{r}
# Find the average income by excluding the missing values
income_avg <- mean(crdata$Income, na.rm=TRUE)

#print average income
print(income_avg)
```

```{r}
# Replace missing values in Income column with average income
crdata_replace <- replace_na(crdata, list(Income = income_avg))
```

```{r}
# Remove records with missing values and assign it to crdata_removed
crdata_removed <- na.omit(crdata_replace)

# Check the summary of the dataset again
summary(crdata_removed)
```

```{r}
# Calculate the number of removed records
nrow(crdata) - nrow(crdata_removed)
```

#### Illustrate the relation between two features; “Income” and “Rating”.
### ggplot
```{r}
# ggplot(data, aes(x, y)) + <geom_function>()
```

```{r}
# Use geom_point(color = "steelblue") to add a scatter plot with blue colour
ggplot(crdata_removed, aes(x = Income, y = Rating)) + geom_point(color= "steelblue")
```