---
title: "3. Information Gain"
date: '2022-06-19'
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
---


```{r}
# install.packages("FSelector")

# Load FSelector package for Feature Selection
library(FSelector)  

# Load "caTools" package for data partitioning
library(caTools)

# Load tidyverse package 
library(tidyverse)

# Import data set and save it as empdata
empdata <- read.csv("EmployeeData.csv", stringsAsFactors = TRUE) #convert string variables to factor variables
```

### check structure
```{r}
# Check the summary of the dataset
summary(empdata)

# Check the structure of the dataset
str(empdata)
```

### redundant variables
```{r}
# Remove redundant variables
empdata[c("EmployeeCount", "EmployeeNumber", "Over18", "StandardHours")] <- NULL
```

### split data
```{r}
# Set a seed
set.seed(10)

# Generate a vector named partition for data partitioning 
partition = sample.split(empdata$Attrition, SplitRatio = 0.8) 

# Create training set: training
training = subset(empdata, partition == TRUE) 

# Create test set: test
test = subset(empdata, partition == FALSE) 
```

### feature selection
### information.gain()
```{r}
# information.gain(target~.,dataset)
```

```{r}
# Use function information.gain to compute information gain values of the attributes
attr_weights <- information.gain(Attrition~. , empdata)

# Print weights
print(attr_weights)
```
### sorting the result
### order()
#### Use order() function to sort the attributes with respect to their information gain values. Then, use barplot() function to illustrate the result.
```{r, fig.align='center'}
# Sort the weights. Use order() function
sorted_weights <- attr_weights[order(attr_weights$attr_importance), ,drop = FALSE]

# Plot the sorted weights
barplot(unlist(sorted_weights), 
        names.arg = rownames(sorted_weights), las = "2", cex.names=0.7,
        ylim = c(0,0.05), space = 0.5)
```

```{r}
# Filter features where the information gain is not zero
library(dplyr)
attr_weights %>% filter(attr_importance > 0)
```

### cutoff.k()
####  filter the most informative k attributes
#### cutoff.k() orders the attributes according to their information gain and returns the first k.
#### cutoff.k.percent(weights, k) selects k* 100% of attributes.
#### cutoff.biggest.diff(weights) selects a subset of attributes which are significantly better than others.
```{r}
# cutoff.k(weights,k) 
```

```{r}
# Use cutoff.k() to find the most informative 19 attributes
filtered_attributes <- cutoff.k(attr_weights, 19)

# Print filtered attributes
print(filtered_attributes)

# Use cutoff.biggest.diff() to a subset of attributes which are significantly better than other
cutoff.biggest.diff(attr_weights)
```

### ggplot
#### plot “Attrition” vs “OverTime”
```{r, fig.align='center'}
library(tidyverse)
ggplot(empdata, 
      aes(x = Attrition, group = OverTime)) + 
      geom_bar(aes(y = ..prop.., fill = factor(..x..)), 
                   stat="count", 
                   alpha = 0.7) +
      geom_text(aes(label = scales::percent(..prop..), y = ..prop.. ), 
                   stat= "count", 
                   vjust = -.1) +
      labs(y = "Percentage") +
      facet_grid(~OverTime) +
      scale_fill_manual("Attrition" ,values = c("steelblue","orange"), labels=c("No", "Yes")) + 
      theme(plot.title = element_text(hjust = 0.5)) + 
      ggtitle("Over time")
```

### rename categories
### revalue()
```{r}
# Revalue categories for the plot. Load 'plyr' package
library(plyr)

# Rename categories for illustration
training$JobRole <- revalue(training$JobRole, 
                            c("Healthcare Representative" = "HealthRep", 
                              "Human Resources" = "HR",
                              "Laboratory Technician" = "LabTech",
                              "Manager" = "Mgr",
                              "Manufacturing Director" = "ManufDir",
                              "Research Director" = "ResDir",
                              "Research Scientist" = "ResSci",
                              "Sales Executive" = "SaleExec",
                              "Sales Representative" = "SaleRep"))

barplotdata = table(training$Attrition, training$JobRole)

# Use barplot function to plot Attrition vs JobRole
barplot(barplotdata, main = "Attrition vs Job Role",
        xlab="JobRole",col=c("steelblue","orange"),
        legend=rownames(barplotdata), cex.names = 0.70, beside = TRUE)
```
 
### plot Attrition vs JobLevel
```{r}
# Plot Attrition vs JobLevel
barplotdata2 = table(training$Attrition, training$JobLevel)


barplot(barplotdata2,
        main="Attrition vs JobLevel",
        xlab="Job Level", col=c("#386cb0","#fdb462"),
        legend=rownames(barplotdata), cex.names = 0.75, beside = TRUE)
```
### plot Attrition vs Monthly Income
```{r}
# Plot Attrition vs Monthly Income
ggplot(training, aes(x = MonthlyIncome, fill = Attrition)) + 
  geom_density(alpha = 0.7) + 
  scale_fill_manual(values = c("#386cb0","#fdb462"))
```
### subset training set
```{r}
# Select a subset of the dataset by using filtered_attributes 
datamodelling <- training[filtered_attributes]
```

#### Since filtered_attributes does not include `the target variable`, `Attrition` column is not present in our constructed data file. Adding it to the data file is needed for model building .
```{r}
datamodelling["target"] <- training["Attrition"]
# or
datamodelling$target <- training$Attrition
```